
Outcomes:

- Create data visualizations using ggplot2
- Explain the idea of declarative graphics

References:

- [R for Data Science Visualization chapter](https://r4ds.had.co.nz/data-visualisation.html)


## ggplot2 philosophy


Structure of data is **vital** for ggplot2.
Your data must be a data frame, with variables in columns- that's how it works.


## Examples

Code from template

```{r}
even = function(x)
{
    win = (x %% 2 == 0) & (x != 0)
    ifelse(win, 1, -1)
}


# only accepts 1 argument, x
column1 = function(x)
{
    ifelse(x %% 3 == 1, 2, -1)
}


# accepts two arguments, x and d
# d has a default of 1.
straightup = function(x, d = 1)   # <--
{
    ifelse(x == d, 35, -1)
}


# Construct a simple betting strategy
simple_strategy = function(bet = even)
{
    function(x) cumsum(bet(x))
}


#' Simulate plays from a betting strategy
#'
#' @param strategy
#' @param nplayers number of players to use this strategy
#' @param ntimes number of times each player should play
play = function(strategy = simple_strategy()
    , nplayers = 100L
    , ntimes = 1000L
    , ballvalues = 0:36
){
    out = replicate(nplayers
            , strategy(sample(ballvalues, size = ntimes, replace = TRUE))
            , simplify = FALSE
    )
    data.frame(winnings = do.call(c, out)
            , player = rep(seq(nplayers), each = ntimes)
            , time = rep(seq(ntimes), times = nplayers)
    )
}
```


## loading a package

Now we can get into the plotting code.
Before we can use ggplot2 we need to load the package with `library`.

```{r}
#install.packages("ggplot2")
library(ggplot2)
theme_set(theme_bw())
```

## generate data

```{r}
class(even)
```


```{r}
set.seed(8930)
TIMES = 20L
d = play(strategy = simple_strategy(even)
         , nplayers = 1, ntimes = TIMES)
head(d)
```


## building a plot

```{r}
g = ggplot(data = d, mapping = aes(x = time, y = winnings))
class(g)
```

```{r}
print(g)
```


123 GO - what element are we missing?
Three elements are required for any ggplot:

1. data
2. mapping
3. geom


## `+` at end of line

```{r}
g = ggplot(data = d, mapping = aes(x = time, y = winnings)) + 
    geom_point()
```


## compare to default `plot`

```{r}
with(d, plot(time, winnings))
```

123 go - which appears to be simpler?

`geom_step`

```{r}
g = ggplot(data = d, mapping = aes(x = time, y = winnings)) + 
    geom_step()  # example of a high level graphical description
```


## More data

Suppose we want to compare multiple betting strategies by drawing two lines on the same plot.
We need more data.

```{r}
dc = play(strategy = simple_strategy(column1)
          , nplayers = 1, ntimes = 30)

```

Let's add a column describing the strategy.

```{r}
d$strategy = "even"
dc$strategy = "column"

# stack d and dc on top of each other
d2 = rbind(d, dc)

# puts d and dc side by side
#d3 = cbind(d, dc)
```


```{r}
a = data.frame(x = 1:3)
b = data.frame(x = 4:6)
rbind(a, b)
cbind(a, b)
```


```{r}
g = ggplot(data = d2, mapping = aes(x = time, y = winnings)) + 
    geom_line()
print(g)
```

Split the lines into groups by strategy, group by the column named `strategy`

```{r}
g = ggplot(data = d2, mapping = aes(x = time, y = winnings)) + 
    geom_line(aes(group = strategy))
print(g)
```

Color the lines according to the values of the `strategy` column.

```{r}
g = ggplot(data = d2, mapping = aes(x = time, y = winnings)) + 
    geom_line(aes(color = strategy))
print(g)
```

`aes` for aesthetic

```{r}
g = ggplot(data = d2, mapping = aes(x = time, y = winnings)) + 
    geom_step(mapping = aes(color = strategy))
print(g)
```






```{r}
r = 0:6
straightup(r, d = 2)
```

Pick random `d` every time.

```{r}
dr = sample(0:36, size = length(r), replace = TRUE)
dr
straightup(r, d = dr)
```

What does `replace = TRUE` do?

```{r}
sample(1:10, size = 20, replace = TRUE)
```

```{r}
sample(1:10, size = 10, replace = FALSE)
```


`d` had 20 rows, `dc` has 30 rows.
```{r}
e = rbind(d, dc, d, d)
```

