Outcomes:

- Identify side effects in function calls
- Predict behavior of reference objects in simple functions

123 GO – Did the student review videos improve your understanding?
Or any comment at all about the videos.

Announcements:

- Turn in last HW (predicting interest) even if you didn't finish


## side effects

Functions take an input and produce an output.
Functions always return a result.
A __side effect__ in a function call is when the function does something - anything at all - in addition to returning the result.

123 GO - Do the following functions have side effects?
Yes, no, not sure

1. `1 + 2`                    No
2a. `hist(1:5)`               Yes - we see the plot
2b. `hist(1:5, plot = FALSE)` No
3. `x = 3`                    Yes
4. `seq(from = 0, to = 10)`   No
5. `read.csv("data.csv)`      No
6. `write.csv(iris, "iris.csv")`    Yes
7. `rnorm(3)`                 Yes


## Pure functions

Pure functions don't have side effects, they always the same output for the same input.
`rnorm` is not pure, because it gives us different outputs.
functions you study in math are pure functions.

why do we like pure functions?

- Easier to reason about.
- parallel programming


## Motivating question

Consider the following minimal reactive Shiny app.
When we run this app, the user inputs a number `a`, and then R generates a message saying what the current value of a is.

Should `ui` and `s` be tied together/ always come in a pair?
Yes.
```{r}
ui <- fluidPage(
  numericInput("a", label = "some number", value = 1)
  , textOutput("msg")
)

s <- function(input, output) {
  
  output$msg = renderText({
    paste("the value of a is", input$a)
  })
  
  NULL
}

shinyApp(ui = ui, server = s)
```

How and why does the function `s` work?

1. What are the arguments to `s`?
      `input`, `output`
2. What does `s` assume about the arguments?
      `output$msg = ...`, `input$a` can be coerced 
3. What does `s` return?
      `NULL`
4. Based on the reactive behavior of the app, what must be happening to the arguments to `s`?
      They're being dynamically updated.
      The arguments to s are __reference objects__
5. Where is `s` called?    `s()`
      Must be called somewhere inside shinyApp
6. Where does the program use `s`?
      Passing `s` as an argument to `shinyApp`
7. Does it make sense to use `s` with `ui` different from the one defined above?


## Reference Semantics

Suppose we have an object `x`.
If a function modifies `x` when it is called, and those changes are visible outside of that particular function call, then `x` has __reference semantics__.
Thus reference semantics are a property of the object, not the function.

Shiny is implemented with reference classes, which have reference semantics, and that's the key to understanding how it all works.
This reference behavior is uncommon, or else at least well hidden, across the rest of the R world.
All the standard objects we work with: vectors, data frames, lists, functions, don't have obvious reference semantics.

We can see reference semantics in action by using environments, and compare the behavior to lists.

```{r}
```


## Plotting on homework - prediction

Standard assumptions:

```{r}
a = 5
b = 10
x = seq(from = 0, to = 1, length.out = 1000)
noise = rnorm(length(x))
y = a*x + b + noise

fit = lm(y ~ x)
```

residuals are the difference between what the model predicts and actual value.

```{r}
r = residuals(fit)
hist(r)
```


```{r}
a = 5
b = 10
x = seq(from = 0, to = 1, length.out = 1000)
noise = runif(length(x))
noise[1:5] = 20
y2 = a*x + b + noise

fit2 = lm(y2 ~ x)
```

Prediction on new data:

```{r}
predict(fit2, data.frame(x = 1:5))
```

residuals are the difference between what the model predicts and actual value.

```{r}
r2 = residuals(fit2)
hist(r2)
```

Separating into 20% test set
`mtcars`

```{r}
n = nrow(mtcars)

test_index = sample(n, size = round(0.2 * n))

test_cars = mtcars[test_index, ]
train_cars = mtcars[-test_index, ]
```









