
Goal:

- Describe the use cases and ways to select subsets of data

Announcements:
Midterm in two parts.

- In class part of midterm through Canvas quizzes on Friday, October 30th.
- Let me know ASAP if you have technical / internet / scheduling issues.
- Multiple choice / true false, etc.
- Low stakes, low pressure.
- Take home part of midterm due Sunday, November 1st.
- Watch the review videos, the questions will be on those concepts.

123 GO- What's your word of the day?

Plan today:

- Review mid quarter check in.
- Showcase some of the unique plots from Q3 in the Roulette homework.
- If time: Group activity to review all the ways to subset vectors and data frame.

Goal today is to step back and review / compare and contrast all the ways we know to select subsets of a data frame.

## Group activity

Come up with all the ways you know how to select subsets of a vector or data frame.
We'll do examples with the `mtcars` data set, because cars are on my mind right now. 😁
We'll take what you come up with and try to make a somewhat comprehensive list of the ways to subset a data frame.

Students to share next time:

- Gao, Garrett, Ben, Hunter, Aleksey

# Vectors


## logical


## index


## names


# Data Frames


## logical


## index


## names



Legends and plots

```{r}
library(ggplot2)
```

```{r}
g = ggplot(mtcars, mapping = aes(x = disp, y = mpg)) +
  geom_point(aes(color = as.factor(cyl)))
print(g)
```

Base R is generally more flexible
```{r}
plot(1:10)
lines(c(2, 8), c(3, 9), lty = 2)
lines(c(2, 8), c(1, 7), lty = 3)
legend("bottomright", legend = c("first", "second", "third"), lty = 1:3)
```
