Outcomes:

- Describe R's type hierarchy
- Predict behavior from implicit coercion

Announcements:

- Any questions on midterm review video assignment?
- I've been grading last homework, should have it back to you this weekend.
- Stay after class if you want to talk cars with me.

123 GO - favorite video game?


## classes and types

The `class` of an object describes how it behaves.
The `type` of an object describes how it is represented.
Sometimes they're the same in R.

```{r}
d = Sys.Date()
d
```

```{r}
class(d)
weekdays(d)
```

```{r}
typeof(d)
```


## doubles

What is a "double"?
Number with decimals, double precision floating point number.

```{r}
dn = as.numeric(d)

weekdays(d)

#123 GO - does this make any sense?
weekdays(dn)

class(dn)
typeof(dn)
```

```{r}
typeof(pi)
```

double precision numbers APPROXIMATE the Real numbers.
They represent numbers using 64 bits.

This implies:
- only finitely many. Can't be more than `2^64` distinct numbers.

epsilon- arbitrarily small number that's still greater than 0.

```{r}
.Machine
```
```{r}
.Machine$double.eps
```

```{r}
1 / (2^(53-1))
```


## explicit coercion

```{r}
paste("hello", "world")
```

Explicit coercion using as.character
```{r}
dc = as.character(d)

typeof(d)
typeof(dc)
```

```{r}
paste("hello", dc)
```


## implicit coercion

123 GO - does this make sense?

```{r}
paste("hello", d)
```


## type hierarchy

Try to order these in terms of when it makes sense to coerce one type to another.

123 GO - order from most specific, to most general

Common types for vectors:

- logical `TRUE`
- integer `seq(5)`, `3L`
- numeric (double) `2.83`
- character `"hello"`

When we combine two vectors using `c`, what type of vector do we get?


## Group activity

Alphabetically, first person's turn to explain where you're at with understanding.
No expectation to understand fully!

1. Explain what is the type of a literal `NA`, and why?

NA starts as a logical and is coerced to more general types according to the type according to the type hierarchy 

```{r}
typeof(NA)
```

```{r}
nai = c(1L, NA)
nai[2]
typeof(nai[2])
```

```{r}
nad = c(1.0, NA)
nad[2]
typeof(nad[2])
```

```{r}
c(1L, as.character(NA))
```

2. Why are these numbers not equal?
```{r}
(0.1 + 0.05) == 0.15
a = 0.1 + 0.05
```

There's a very small difference because of the way floating point math works.

```{r}
a - 0.15
```


2b. Why are these numbers not equal?
```{r}
pi == as.numeric(as.character(pi))
```

3. `identical` checks if objects are exactly the same, in every way.
Explain the results of these function calls:
```{r}
1 == 1.0
1 == 1L
identical(1, 1.0)
identical(1, 1L)
```

4. What is the type of an object with class `data.frame`, and what does this imply?

```{r}
typeof(iris)
```

```{r}
l = list(1:3, 7)
```

We can call functions that expect lists as inputs

```{r}
lapply(iris, class)
```
