
Goal:

- Describe the use cases and ways to select subsets of data

Announcements:

- Take home part of midterm posted
- Friday (next class) we'll do in class part of midterm

123 GO- What's your animal of the day?

Plan today:

- Review subset data group activity
- 2nd person in group: Gao, Garrett, Ben, Hunter, Aleksey
- Demonstrate midterm format


## Group activity

Come up with all the ways you know how to select subsets of a vector or data frame.
We'll do examples with the `mtcars` data set, because cars are on my mind right now. 😁
We'll take what you come up with and try to make a somewhat comprehensive list of the ways to subset a data frame.

In summary the ways to subset are:

- logical, define a condition
- index, integers for position
- name, using character vector of names

123 GO- which one doesn't allow repetition of elements?
Logical



# Vectors

```{r}
x = seq(from = 0, to = 1, by = 0.1)
x
```

## logical

```{r}
x[x < 0.5]
```

## index

```{r}
x[c(1, 9, 1, 3)]
```

## names

123 GO: Can I index `x` with names right now?
No.

```{r}
names(x)

seq_along(x)

letters[seq_along(x)]

names(x) = letters[seq_along(x)]
x
```

```{r}
x[c("g", "g", "a")]
```


# Data Frames

```{r}
head(mtcars)
```

General form:
```
DATA[rows, columns]
```

## logical

Super useful.

rows with `mpg` greater than 20.

```{r}
mtcars[20 < mtcars$mpg, ]
```
```{r}
l = 20 < mtcars$mpg
mtcars[l, ]
```

`with` function

```{r}
with(mtcars, mtcars[20 < mpg, ])
```

`subset` function

```{r}
subset(mtcars, 20 < mpg)
```

## index

```{r}
mtcars[1:10, ]
```


Random sample of rows:
```{r}
mtcars[sample(nrow(mtcars), size = 3), ]
```

Repetition is allowed:
```{r}
mtcars[c(1, 2, 1, 9, 1), ]
```

First 5 rows, first 3 columns:
```{r}
mtcars[1:5, 1:3]
```
Negative indices.
Says what to drop.

```{r}
mtcars[-(1:2), -(3:11)]
```

## names

Select the column `mpg`

`[` is preferred for programming: 

```{r}
mtcars[, "mpg"]
```

Didn't work:
```{r}
mtcars[, "mp"]
```

Convenient for interactive use:

```{r}
mtcars$mpg
```

123 GO: Will the line below work?
Yes, it did work.

```{r}
mtcars$mp
```

To select multiple columns by name:

```{r}
mtcars[, c("mpg", "cyl")]
```

Can pick rows by names:

```{r}
mtcars[c("Valiant", "Mazda RX4"), c("hp", "cyl")]
```

Can combine logical row selection with column selection by name.

```{r}
mtcars[mtcars$mpg > 20, c("hp", "cyl")]
```

## splitting

In general, your code should be as direct and readable as possible.
123 GO, which way is more direct? Most people say mt4.

```{r}
mt2 = split(mtcars, mtcars$cyl)
mt3 = mt2[["8"]]

mt4 = mtcars[mtcars$cyl == 8, ]
```
mt3 is the same subset as mt4.






