123 GO - what's your word of the day?

- Write reproducible data analyses
- Maintain consistency between written code and global environment
- Ask precise questions about programming
- Predict when variables will be available in function bodies and global environment

Announcements:

- Office hours catered to students new to programming
  Fill this out right now!! https://www.when2meet.com/?9867200-mw54z
- Glad to see more code and chatter in the Discord.
    It's good to post your broken code here, so we can fix it.


Resources:

- Stack Overflow on [minimal reproducible example](https://stackoverflow.com/help/minimal-reproducible-example)

[live notes](https://github.com/clarkfitzg/stat128/blob/master/2020-09-21.Rmd)


## Sharing workflow tips in RStudio

Workflow is __how__ you develop your programs.
Workflow can make the difference between a pleasant experience and endless frustration.

How are you currently using RStudio?

In your small groups come up with:

1. One workflow tip that works
    1. Use Youtube
    2. Trial and error
    3. set your working directory, save all your stuff in one folder.
    4. Clear global environment
    5. print tables to make sure they look right!!!! <----
    6. test code in blocks / chunks
    7. use book as reference
    8. write the code piece by piece, so that it always works
2. One place where you have trouble with your workflow.
    1. Need to learn shortcuts
    2. Double check facts and assumptions (Only 30 days in June)
    3. Clear global environment - be careful
    4. Help documentation often not helpful!
    5. understanding exactly what output to expect from a function
    8. knowing when assignment to global environment will happen

```{r}
d = 10
```

```{r}
d = "abc"
```


```{r}
d + 1
```


## How to ask a precise question

When asking questions, copy and paste the code and output that gives you trouble.
Not a screenshot, unless you're asking about plots.
For example...

Example should be minimal, complete, and reproducible.


## Reproducible data analyses

Q: When you go to produce your final report, should it matter what you have in your global environment?

123 GO - yes or no
__NO__ 


Q: Can you think of anything that should not be contained in the code, but rather kept private?



## Function Scope



### Globals

