Outcomes:

- Reason about vectorized boolean computations
- Explain difference between `TRUE` and `T`.

Announcements:

- You should be proud of yourselves for the last HW.
- New programmer specific office hours after class today, all are welcome.

123 GO - What's your word of the day?

## boolean logic

```{r}
TRUE

FALSE
```


### Negation

`!x` does the logical negation of `x`.

```{r}
trues = c(TRUE, TRUE)
falses = !trues
!falses

tf = c(TRUE, FALSE)
```


### AND OR


- `A & B` returns TRUE when A and B are both TRUE, and FALSE otherwise.
- `A | B` returns TRUE if any of A or B are TRUE, and FALSE otherwise.

AND `&`
```{r}
tf & trues
```

123GO: 

```{r}
tf & falses
```

`|` logical OR

```{r}
tf | trues
```

For any value of `x` the following expression will always be `TRUE`:

```{r}
x = sample(tf, size = 300, replace = TRUE)
y = x | TRUE
all(y)
```


### Binary Operators

Side note:
`|` and `&` are examples of __binary operators__.

Other binary operations include addition `+`, and multiplication, `*`.
You can think of them as functions that take two arguments, and are usually called without parentheses in the syntax of a language as `x OP y`, rather than `OP(x, y)`.

```{r}

#1
`&`(tf, trues)

# Both do the same thing. Which is more readable?

#2
tf & trues
```

Logical negation using `!` operates on a single argument, so it's a __unary operator__.

`!` sometimes called NOT, "bang"

```{r}
!tf
```


## Difference between `TRUE` and `T`

```{r}
class(TRUE)

class(T)

identical(T, TRUE)
```

You write simulation involving time.
So you start at time 
```{r}
T = 0
...
```

```{r}
tf & TRUE
```

```{r}
tf & T
```

The issue is that `T` is a variable name, that happens to be assigned to TRUE.
`TRUE` is a string literal constant.

```{r}
TRUE = 0
```

```{r}
?Reserved
```

Moral of the story: Always write `TRUE`.

## Sequences of AND's and OR's

123 GO
What will this return?

```{r}
TRUE | TRUE & FALSE
```

I don't know, honestly.

What can happen?

If `|` is evaluated first.
```{r}
(TRUE | TRUE) & FALSE  # TRUE & FALSE which is FALSE
```

If `&` is evaluated first.
```{r}
TRUE | (TRUE & FALSE) # TRUE | FALSE which is TRUE
```

This implies that `&` happened first, before `|`.

```{r}
?Syntax
```

### Operator precedence



Best practices:

- Use parentheses to clarify.
    That's what we did with `(TRUE | TRUE) & FALSE`.
- Combine several conditions with exactly one of `|` and `&`.
    If you do this, operator precedence doesn't matter.
- Don't mix `&` and `|` in long logical expressions.

```{r}
x = 1:10
x[(x <= 5) & (x != 2) & (1.6 < x) & (x != sqrt(pi))]
```


## Group activity

`xor(x, y)` returns `TRUE` if exactly one of `x` and `y` is `TRUE`.
As a group, use only `!, |, &` to implement a function `xor2` with this behavior.

"Exclusive OR"
```{r}
xor(tf, trues)
xor(tf, falses)

xor2 = function(x, y)
{
  (x + y) == 1
}
```


## homework hints

I gave you this function in your homework template to illustrate some patterns you can use.

`%%` called mod for modular arithmetic.
`x %% y` is remainder when `x` is divided by `y`.

```{r}
even = function(x)
{
    win = (x %% 2 == 0) & (x != 0)
    ifelse(win, 1, -1)
}
```

To show this works

```{r}
s = 0:36
even(s)
```


```{r}
iseven = function(x) (x %% 2) == 0
iseven(1:10)
```

`ifelse` is vectorized, operating elementwise.

Expected value

```{r}
s = 0:36
```

For discrete RV with k possible outcomes
$$
E[X] = \sum_{i=1}^k p(X = x_i) * x_i
$$

Example:

```{r}
tb = table(even(s))

-1 * (19/37) + 1 * (18/37)
```

Because (19/37) is the probability that this discrete random variable with two outcomes takes on the value -1.

Manipulating the table:

```{r}
tb
class(tb)
```

Suppose I want to extract 19 and 18

```{r}
as.numeric(tb)
```

Names of the table:

```{r}
class(names(tb))

vals = as.numeric(names(tb))
```


```{r}
x = 100
f = function(x)
{
  x + 1
}
f(1)

f(x)

# First x refers to the one in the function body, the second to the global x
f(x = x)
```


```{r}
# group these together at the top of your program, it's easier to read.
LINETYPE = "l"
COLOR = "blue"

plot_wrapper = function(x, y, linetype = LINETYPE)
{
  plot(x, y, type = linetype)
}

plot_wrapper(1:10, 1:10)
```

first dozen bet

Corner cases: 0, 1, 12, 13, ...

```{r}
f(0) == -1
f(1) == 2
f(12) == 2
f(13) == 2
```


$$
f(x) =  \begin{cases}
-1 &\mbox{if } x > 12 \\
2 & \mbox{if } x \leq 12
\end{cases}
$$
$$


```{r}
```


Rocky's prime bet function:

```{r}
# This works on scalars
is.prime0 <- function(n) n == 2L || all(n %% 2L:max(2,floor(sqrt(n))) != 0)

is.prime <- function(n)
{
  ipv = Vectorize(is.prime0)  # A bit of a hack
  if(1 < length(n)){
    ipv(n)
  } else {
    is.prime0(n)
  }
}

d = data.frame(n = 0:36, prime = is.prime(0:36))


prime = function(x)
{
  win = is.prime(x) == TRUE
  ifelse(win,1,-1)
}
```


The amount you win or lose.
Place $10 on a bet that pays 5:1.
If you win: you get 5 * (bet amount) = 5 * 10 = $50.
If you lose: you lose $10.

Playing roulette:

Bet $1 on lows.
Ball hits the 22.
Then you lose $1.

Bet $1 on lows.
Ball hits the 12.
Then you win $1.

Bet $1 on first dozen, with payout 2:1.
Ball hits the 12.
Then you win $2.

Anytime you bet on the first dozen, you either lose $1, or you win $2.

```{r}
ifelse(win_firstdozen, 2, -1)
```


