Goals:

- Write wrapper functions
- Critique different styles of code

123 GO – What's one item on your Christmas wish list?


## Wrapper functions

Suppose I want to take an existing function and modify it slightly, say by giving the function arguments different defaults.
For example, `pnorm` calculates the cumulative distribution function for the normal distribution:

```{r}
x = seq(from = -4, to = 4, by = 0.1)

plot(x, pnorm(x))

lines(x, pnorm(x, sd = 2))
lines(x, pnorm2(x), col = "red")
```


I want to create a function `pnorm2` that behaves just like `pnorm`, but has the default standard deviation (sd) set to 2.
In other words, `pnorm2(x)` should behave exactly like `pnorm(x, sd = 2)`.
How can I do this?

```{r}
pnorm2 = function(x)
{
    pnorm(x, sd = 2)
}
```

123 GO: can we change sd in a call to pnorm2?
No- 2 is hardcoded in, can't change it.

123 GO: Is there any point in having pnorm2, if we have pnorm3?
No, pnorm3 is more flexible.

123 GO: Is it possible to specify `lower.tail = FALSE` in a call to pnorm3?
No.

```{r}
pnorm3 = function(x, sd = 2)
{
    pnorm(x, sd = sd)
}

plot(x, pnorm3(x))
```

```{r}
pnorm4 = function(x, sd = 2, lower.tail = TRUE)
{
    pnorm(x, sd = sd, lower.tail = lower.tail)
}

plot(x, pnorm4(x))
```

What happens when `pnorm` gets a *new* argument, `fast = TRUE`.
123GO: Is it possible to use this new argument?
No. Not without changing `pnorm5`.

General principle for programming: Do as little as possible.

```{r}
pnorm5 = function(x, mean = 0, sd = 2, lower.tail = TRUE, log.p = FALSE, fast = TRUE)
{
    pnorm(x, mean = mean, sd = sd, lower.tail = lower.tail, log.p = log.p)
}

plot(x, pnorm5(x))
```

`...` called ellipses.
"Syntactic sugar" pretty way to express something in a programming language.

```{r}
pnorm6 = function(..., sd = 2)
{
    pnorm(..., sd = sd) # <--- 
}

plot(x, pnorm6(x, lower.tail = FALSE))
```

```{r}
pnorm7 = function(..., sd = 2)
{
    args = list(...)
    #browser() # STOP the R interpreter right here, so that I can look around
    print(paste("mean is ", args$mean))
    pnorm(..., sd = sd) # <--- 
}

pnorm7(x, lower.tail = FALSE, mean = 3)
```


## Group activity

10 minutes

1. Write a wrapper for `hist` that changes the default title (parameter `main`) to "HISTOGRAM".
2. List the differences between the functions `extract_interest_rate` and `eir` in https://github.com/clarkfitzg/stat128/blob/master/extract_interest.R
Which do you prefer, and why?


```{r}
#' Extract interest rates from a character vector
#'
#' @param x character vector containing text with interest rates
#' @param pattern regular expression pattern to find the matching interest rates
#' @return The first numeric interest rate found in the string, or NA if none found
#' @examples
#' extract_interest_rate("Fixed @ 3.92%, no TIC/NIC")  # 3.92
#' extract_interest_rate("123.456 but no adjacent %")  # NA
extract_interest_rate = function(x, pattern = "[0-9]+\\.[0-9]+\\%")
{
    interest = rep(NA, length(x))
    match_locations = regexpr(pattern, x)
    match_string = regmatches(x, match_locations)
    match_number_string = gsub("\\%", "", match_string)
    match_number = as.numeric(match_number_string)
    interest[match_locations != -1] = match_number
    interest
}


eir = function(x, p = "[0-9]+\\.[0-9]+\\%")
{
    i = rep(NA, length(x))
    l = regexpr(p, x)
    n = as.numeric(gsub("\\%", "", regmatches(x, l)))
    (i[l != -1] = n)
}
```

What are the differences between these functions?

Which version of this function is easier to read?

Which version will be easier to debug?