Outcomes:

- Merge two data frames that share a common column

123 GO – What's your word of the day?

Announcements:

- Submit final project highlight separately from final project on Sunday.
- Evaluate this course.

To review:

- barplots ggplot `+ geom_bar`

```{r}
library(ggplot2)

people = read.csv(text = "name,country,height
Clark,US,73
Jen,US,65
Sara,Mexico,54
Antonio,Honduras,56
Alexander,Russia,70
Sofia,Russia,68")

#g = ggplot(data = people, mapping = aes(x = name, y = height)) + geom_bar()
#g = ggplot(data = people, mapping = aes(x = name, y = height)) + geom_bar(stat = identity)
g = ggplot(data = people, mapping = aes(x = name, y = height)) + geom_col(aes(fill = country))
#g = ggplot(data = people, mapping = aes(x = name, y = height)) + geom_point() # Works
print(g)
```


- constructing data frames from csv file `read.csv`

```{r}
# Round trip from data frame -> CSV -> data frame
getwd()
#write.csv(iris, "iris.csv")
i2 = read.csv("iris.csv")
i3 = read.csv("iris.csv", skip = 4)
```


## Merging

Suppose I know which country a person is from, and I want to greet them using a greeting based on the language of their country.
If we had a column representing the greeting for each person, then we could just use `paste`.

```{r}

country = read.csv(text = "country,language
US,English
Russia,Russian
Mexico,Spanish
Honduras,Spanish")

greeting = read.csv(text = "language,greeting
English,hello
Spanish,hola
Korean,안녕
Russian,Здравствуйте")
```

merge country and greeting on the language column
merge people with country on the country column

```{r}
country_greeting = merge(country, greeting)

pcg = merge(people, country_greeting)

paste("Clark", "hello")

# 123 GO: what will this produce?
paste(pcg$name, "hello")

paste(pcg$name, pcg$greeting)
```


## Nested data

JSON, XML

```{r}
x = list(a = 1:5, b = "hello",
         c = list(6:10, list(letters[2:4], pi))
         )
```

## Defining a new vector

based on the output of `grepl`.

```{r}

input = strsplit("Hillary campaign fund, Trump campaign fund, Romney campaign fund, Sanders campaign fund, Hillary and Trump weekend getaway", split = ", ")[[1]]

output = rep("Other", length(input))

for(term in c("Hillary", "Trump", "Romney")){
    found = grepl(term, input)
    output[found] = term
}

output
# [1] "Hillary" "Trump"   "Romney"  "Other"   "Trump"


```


