
- Write functions with `for` and `while` loops
- Know when to write loops, and when to use vectorized functions or `lapply` and related functions

Announcements:

- Midterm review video assignment is live.

123 GO - What movie character would you like to meet?


## `for` loops

`for` loops are used to iterate through vectors.

```{r}
for (i in 1:5) {
    print(i^2)
}
```

```
for (variable in vector) {
    #RUN THIS CODE
}
```

```{r}
variable = vector[[1]]
# RUN THIS CODE

variable = vector[[2]]
# RUN THIS CODE

...

variable = vector[[length(vector)]]
# RUN THIS CODE
```



## Alternatives to `for` loops

```{r}
x = c(10, 2, 1, 7)

# Say you want to take log of every element of x
ans = rep(NA, length(x)) # Preallocation
for(i in seq_along(x)){
    ans[i] = log(x[i])
}
```

We could use an `apply` function.

Survey (123 GO), which is more clear? The loop, or sapply?

```{r}
sapply(x, log)
```

Vectorization loops over each element and applies the function.
Preferred idiomatic R code, so use it whenever possible.

```{r}
log(x)
```



## When to use loops

Use explicit loops when there is true loop carried data dependence.
Meaning- what you do in one iteration affects what you do in a future iteration.

123 GO
What is the argument for the function?
Suppose I want to calculate the nth fibonacci number.

```{r}
fib = function(n = 5){
    a = 0
    b = 1
    for(i in seq(n)){
        apb = a + b # Here we use a and b, which were defined in previous iteration
        a = b
        b = apb
    }
    b    
}

fib()
fib(6)
fib(7)
```


## `while` loops

```{r}
n = 0
while(n < 5){
    print(n)     # body of the loop
    n = n + 1
}
```
general form

```{r}
while(CONDITION IS TRUE){
    RUN THIS BLOCK
}
```

Approximately:

```{r}
if(CONDITION IS TRUE){
    RUN THIS BLOCK
} else if(CONDITION IS TRUE){
    RUN THIS BLOCK
} else if(CONDITION IS TRUE){
    ...
    Forever
```


## `break` and return

`break` is used to break out of a loop early.

```{r}
for(i in 1:10){
    print(i)
    if(i == 5){
        break
    }
}
print("all done")
```
123 GO- the last number that we see before "all done"






## Example:

Here's a more compelling example in the context of our roulette martingale example.
Suppose we would like to look at the histogram of the number of times you can play with this strategy until you're broke.

```{r}
times_played_until_broke = function(maxplays = 1e4, money = 5000, initialbet = 10){
    bet = initialbet
    for(i in seq(maxplays)){
        win = sample(c(TRUE, FALSE), size = 1, prob = c(18, 19)/37)
        if(win){
            money = money + bet
            bet = initialbet
        } else {
            money = money - bet
            if(money <= 0){
                # Once we have no money left, there's no need to keep going.
                break
                # Alternatively:
                # return(i)
            }
            bet = 2 * bet
        }
    }
    i
}

times_played_until_broke()

x = replicate(500L, times_played_until_broke())

hist(x)
```

The way we've written this, the value `10,000` actually means "more than 10,000 plays".
We can still look at the quantiles to get an idea of how many times we could play like this:

```{r}
quantile(x)
```

```{r}
#' Suppose you want to **return** NA if the player can play more than 10,000 times
times_played_until_broke2 = function(maxplays = 1e4, money = 5000, initialbet = 10){
    bet = initialbet
    for(i in seq(maxplays)){
        win = sample(c(TRUE, FALSE), size = 1, prob = c(18, 19)/37)
        if(win){
            money = money + bet
            bet = initialbet
        } else {
            money = money - bet
            if(money <= 0){
                # Once we have no money left, there's no need to keep going.
                return(i)
            }
            bet = 2 * bet
        }
    }
    NA
}

x2  = replicate(500, times_played_until_broke2())

hist(x2)
```



Exercise: Change the above function to use a `while` loop.

