
123 GO - what's your word of the day?

- Create and interpret contingency tables
- Create and interpret interaction plots

Announcements:

- Homework is posted.
- Data set is about 400 MB, let me know if your computer starts going super slow.

If your computer is slow:

1. Close other programs, like your web browser
2. Load a subset of the data


## Categorical Data

Categorical data usually has a handful of distinct levels.
For examples, colors.
R uses __factors__ to represent categorical data.

```{r}
colors = c("red", "blue", "green")
class(colors)
c2 = factor(colors)
c2
```

```{r}
ed = c("HS", "BA", "MS", "PHD")
edf = factor(ed)
```

`factor` constructs factors from character vectors.

Sometimes we have ordered factors.

It makes sense to compare ordered factors.

```{r}
edo = factor(ed, levels = ed, ordered = TRUE)

edo[1] < edo[4]

# Not meaningful
# edf[1] < edf[4]
```

Suppose we would like to make the continous data in x into a factor.
How can we do this?

```{r}
x = seq(from = 1.5, to = 10, length.out = 100)
xf = factor(x)
x2 = as.numeric(xf)
x3 = as.numeric(as.character(xf))
```

Side note: be very careful with equality comparison for floating point numbers.

```{r}
# Incorrect
head(x) == head(x3)

# Right way
epsilon = 1e-14
abs(head(x) - head(x3)) < epsilon
```

```{r}
cx = cut(x, breaks = 1:10)
```


## Contingency tables

```{r}
# Default on a mac
covid = read.csv("~/Downloads/COVID-19_Case_Surveillance_Public_Use_Data.csv")

# I use "~/data" as my directory, because that's where I keep large data sets.
covid = read.csv("~/data/COVID-19_Case_Surveillance_Public_Use_Data.csv")
```

What kind of data is `age_group`?
Is it a factor? - 123GO
Conceptually, yes, it is a factor

```{r}
table(covid$age_group)
head(covid$age_group)

class(covid$age_group)
```

```{r}
table(covid$medcond_yn)
```

Suppose you want to look at a table with both age group and medical condition.

Contingency table:
```{r}
tam = table(covid[, c("age_group", "sex")])

plot(tam)
```


## Response and predictors

Many models in statistics and machine learning are based on the idea of using one or more variables to predict a different variable of interest, called the __response__.

Hospitalization

```{r}
table(covid$hosp_yn)
class(covid$hosp_yn)
covid$hospital = as.numeric(covid$hosp_yn == "Yes")
```

Goal: model hospitalization as a function of age and sex

```{r}
with(covid,
     interaction.plot(age_group, sex, hospital)
     )
```


Severity of covid 19:

- 0 no hospitalization
- 1 hospital
- 2 ICU
- 3 died

## Interaction plots

In statistics, an __interaction effect__ is when two or more variables have a non additive effect in the value of some response.



## Group Activity

Discuss in your groups, and come up with examples of:

1) unordered categorical variable
2) ordered categorical variable
3) two variables that may well have an interaction effect in producing a response
