Goals:

- Simulate data from R's builtin distributions  prefix `r<...>` 
- Calculate and graph probability density / mass functions  `d<...>`

123 GO – Which class are you most looking forward to next semester?

Announcements:

- This is last homework - Shiny
- Have early check in on the project

References:

CRAN views- Distributions

- https://cran.r-project.org/web/views/Distributions.html


## Stand on the shoulders of giants

If X ~ Poisson(140), then the probability mass function is:

$$
P(X = k) = e^{-140} \frac{140^k}{k!}
$$

$$
P(X = k) = e^{-\lambda} \frac{\lambda^k}{k!}
$$

Think of Poisson as modeling counts, where the upper bound is unlimited.

We want to calculate P(X = 3)

```{r}
pmf140 = function(k, lambda = 140)
{
  exp(-lambda) * lambda^k / factorial(k)
}

pmf140(3)
```

```{r}
exp(-140)
```

```{r}
140^c(143, 144)
```

```{r}
k = 130:143
y = pmf140(k)
plot(k, y, ylab = "probability mass function")
```

```{r}
pmf140(135:143)

pmf140(135:145)
```

The better way is to use the builtin functions in R

```{r}
pmf = dpois(135:145, lambda = 140)
pmf
plot(pmf)
```
```{r}
# Roll yer own
pmf140 = function(k, lambda = 140)
{
  exp(-lambda) * lambda^k / factorial(k)
}

# builtin
dpois(135:145, lambda = 140)
```

123 go: Why is the builtin way preferred?

The builtin functions have been battle tested and refined for decades

1. Clarity - 
2. Less code - easier, less opportunity for mistakes
2. convenience
3. Robust 
3. faster



## R probability function naming conventions

Base R has 4 different probability functions for 16 different distributions, and external packages on CRAN have many more.
The behavior of the function comes from the prefix.

```{r}
# prefix suffix

# to generate random data from a normal distribution
x = rnorm(100)
hist(x, freq = FALSE)

xx = seq(from = -3, to = 3, length.out = 100)
dx = dnorm(xx)
lines(xx, dx, col = "blue", lwd = 3)

```

Prefixes:

- `d` probability density / mass functions
- `p` probability (cumulative) distribution functions
- `q` quantile functions
- `r` random number generation

The distribution comes from the suffix.

Suffixes:

1. `beta` beta
1. `binom` binomial 
1. `cauchy` Cauchy 
1. `chisq` chi-squared 
1. `exp` exponential 
1. `f` Fisher F
1. `gamma` gamma 
1. `geom` geometric 
1. `hyper` hypergeometric 
1. `logis` logistic 
1. `lnorm` lognormal 
1. `nbinom` negative binomial 
1. `norm` normal 
1. `pois` Poisson 
1. `t` Student's t
1. `unif` uniform 
1. `weibull` Weibull 


```{r}
# prefix suffix

# to generate random data from a normal distribution
# r is the prefix
# suffix was norm -> t
x = rt(100, df = 10)
hist(x, freq = FALSE)

xx = seq(from = -3, to = 3, length.out = 100)
# d is the prefix
# suffix was norm -> t
dxnorm = dnorm(xx)
dx = dt(xx, df = 10)
lines(xx, dx, col = "blue", lwd = 3)  #lwd line width
lines(xx, dxnorm, col = "red", lwd = 2)

```

## Quantiles


```{r}
q = seq(from = 0.1, to = 0.9, by = 0.1)
qweibull(q, 0.1)
```

```{r}
x = seq(from = 0, to = 2, length.out = 200)
pdf = dweibull(x, shape = 0.1)
plot(x, pdf, type = "l")
```






