```{r}
library(ggplot2)
```

Outcomes:

- Use `::` and `:::` to find objects in a namespace
- Predict which object will be found based on search paths
- object oriented programming - Define new methods for generic functions like `plot`

123 GO – What are you going to do when finals are over?

Announcements:

- next class Q&A on final project and review for final
    Bring questions.
- would you be interested in working with a real company on a NodeJS / SQL project?
    would you pay $50 for such an experience?
    [Riipen](https://www.riipen.com/)


## The search path

When we start a fresh R session, there are no objects in our global workspace.
We can check with `ls`, list objects.

```{r}
ls()
x = 10
```

R finds objects by looking through packages in a particular order.
That order is called the __search path__, and we can inspect it with `search()`.

```{r}
search()
```

Every time R encounters an object name it looks through the search path to find that object.


Suppose we define our own `ls` function.

```{r}
ls = function(...){
    "Hmmm- is this a good idea?"
}
```

123 GO: What will happen?

```{r}
ls()
```


We can use `::`, the namespace operator, to pick out an object from a namespace.

```{r}
# package::object_name
base::ls()

ls = base::ls
ls()
```


123 GO: What do you think will happen to my search path after attaching a package with `library`?

```{r}
search()
library(ggplot2)
library(MASS)
search()
```

How can an "object not found" error happen?



## How does `plot` work?

Set up:

```{r}
x = 1:10
y = x + rnorm(length(x))
h = hist(x, plot = FALSE)
fit = lm(y ~ x)
```

The function `plot` behaves differently depending on what kind of arguments we pass into it.

```{r}
par(mfrow = c(2, 2))
plot(x)
plot(x, y)
plot(h)
plot(fit, which = 1)
```

What's going on?
There must be a lot of code in `plot` to handle all of these different cases, right?
We can look at it, just print the function out:

```{r}

```


S3 object oriented system
S4 is a more sophisticated, safter system

The entire `plot` function is just a single function call: `UseMethod("plot")`.
`plot` is called a __generic__ function.
The actual implementation happens in the plotting methods, namely:

- `plot.default`
- `graphics::plot.function`
- `graphics:::plot.histogram`
- `stats:::plot.lm`

```{r}
class(x)
class(h)
class(fit)
```

In contrast to `::`, the `:::` "triple colon" picks out objects which are not exported from packages.
These objects are not exported by design- everything works fine without them.
You shouldn't use `:::` in your own code, but it's useful to understand where the objects are, and for debugging.

Let's look at the class of `x`, `h`, and `fit` to see what's going on.

```{r}
class(x)
class(h)
class(fit)
```

R's S3 object oriented system simply dispatches on the class of the object.
For example:

```{r}
plot(h)

graphics:::plot.histogram(h)
```

In other words, `plot` does the right thing when we pass in an object with class `histogram`.
This is one style of object oriented programming.

123 GO: Why is this behavior useful?

- easier to use because you don't need to remember many function names
- Makes for simpler code
- Dynamic coding- same generic function can work and do the right thing for many different arguments


## Extending

We can extend this system by defining our own methods.
For example, suppose we want to attach units to a numeric vector, say seconds.
There are packages to do this, so this is just an example.

Here's a direct way to create an object with a new class.
Usually, we would write a constructor function, but one step at a time.

```{r}
x = 1:10
class(x) = c("units", class(x))
attr(x, "units") = "seconds"
class(x)

attr(x, "units")
```


Our method definition sets the x label based on the units, and then calls the next method.


```{r}
plot.units = function(x, y, xlab = attr(x, "units"), ...){
   # xlab = attr(x, "units")
   NextMethod(xlab = xlab) 
}
```

`NextMethod` calls the next method for `plot` with all the same arguments, except with a new default for `xlab`.
How and why `NextMethod` works is a more advanced topic.

```{r}
y = rnorm(length(x))
plot(x, y, xlab = "SEC")
```

