
- Split data, apply functions, and combine results

Announcements:

- Next assignment will be midterm review video.
- For detailed homework feedback, ask in office hours or Discord
- Data science panel Friday, thanks for submitting questions
- Need 2 people to help me try out webinar feature after class today.
- Homework hint- for the questions on the martingale, it may be easiest to modify the `doublebet` strategy function.
    Ask me on Discord and I'll help you with this, because we haven't covered control flow like loops and conditional statements.

123 GO - What's the last restaurant you ate at?

Resources:

- [The Split-Apply-Combine Strategy for Data Analysis](https://www.jstatsoft.org/article/view/v040i01) by Hadley Wickham

## Review

Last class we looked again at the roulette simulations.

```{r}
source("https://raw.githubusercontent.com/clarkfitzg/stat128/master/roulette.R")

set.seed(1380)
d = play(simple_strategy(even), nplayers = 500L, ntimes = 1000L)
```

Our eventual goal was to compute a couple summary statistics for every value of time in a specific form, so we wrote a function `summary_at_time` to do this for one particular time.

```{r}
#' Returns several summary statistics at a particular time
#'
#' @param dt data frame with only one distinct time value
#' @return output: data frame with one row containing columns mean, sd, and time
summary_at_time = function(dt)
{
    time = unique(dt$time)
    data.frame(time, mean = mean(dt$winnings), sd = sd(dt$winnings))
}
```


```{r}
d1 = d[d$time == 1, ]
summary_at_time(d1)

summary_at_time(d[d$time == 1, ])
```

Goal: do it for all our times.

```{r}
s1 = summary_at_time(d[d$time == 1, ])
s2 = summary_at_time(d[d$time == 2, ])
s3 = summary_at_time(d[d$time == 3, ])
s4 = summary_at_time(d[d$time == 4, ])
s5 = summary_at_time(d[d$time == 5, ])
```
This violates DRY.

Recall DRY: "Don't Repeat Yourself"

```{r}
rbind(s1, s2, s3, s4, s5)
```


How many to do?

```{r}
length(unique(d$time))
```


## `lapply`

"idiom" - a function or pattern used by people who know what they're doing.

`lapply` means "list apply"

```{r}
seq(2)
seq(5)
seq(10)
```

Apply the function `seq` to every value from 1 to 5.

```{r}
l = lapply(1:5, seq)
l
```
123 GO - does it make sense to return this in a data frame?

```{r}
do.call(data.frame, l)
```

```{r}
data.frame(1, 1:2, 1:3)
```


## lists

__list__ is a general purpose data container.

```{r}
l
class(l)
```

Lists can contain anything at all.

```{r}
l[[2]] = head(letters)
l
```

What's good about `lapply` returning a list?
It's the most general kind of object to return, it doesn't make any assumptions about what your function does.


## `split`

We want to split `d`, our big data frame, by the values in `d$time`.

```{r}
ds = split(d, d$time)
```

```{r}
length(ds) == length(unique(d$time))

d539 = ds[[539]]

class(d539)
head(d539)
dim(d539)
```

```{r}
lapply(head(ds), dim)
```

123 GO - what's this?

```{r}
class(ds)
```

Let's apply our function 

```{r}
dr = lapply(ds, summary_at_time)
```

ALWAYS be checking your results

```{r}
head(dr)

s1
s2
s3
```

```{r}
rbind(dr[[1]], dr[[2]], dr[[3]], dr[[4]])
```


## Combining

We want to call `rbind`, using the elements of `dr` as arguments

```{r}
dfinal = do.call(rbind, dr)
```

```{r}
dim(dfinal)

head(dfinal)
tail(dfinal)
```






## Putting it all together

`d` contains all our data.

```{r}
ds = split(d, d$time)
ds2 = lapply(ds, summary_at_time)
dfinal = do.call(rbind, ds2)
```

In the future we'll look at packages data.table and dplyr, they can make this kind of operation simpler and faster.
The purpose of studying it in detail is to learn the concepts, along with some standard R idioms.

Now we can happily create our plot with ggplot2.

```{r}
library(ggplot2)

g = ggplot(data = dfinal, mapping = aes(x = time, y = mean)) +
    geom_line()

print(g)
```


## `lapply` details

`lapply` is roughly equivalent to this code:

`...` is "ellipsis" or "dot-dot-dot" which allows you to pass in more arguments for the function `FUN`.

__higher order function__ is a function that accepts functions as arguments.

```{r}
lapply_idea = function(X, FUN, ...)
{
    result = vector(mode = "list", length = length(X))
    for(i in seq_along(X)){
        result[[i]] = FUN(X[[i]], ...)
    }
    result
}

lapply_idea(1:5, mean, na.rm = TRUE)
```

You might ask, why use `lapply` instead of just a loop then?
There are a few answers:

1) `lapply` uses less code than the loop, which means less opportunities for bugs.
2) `lapply` is fast and idiomatic.
3) `lapply` is easy to make faster through parallel programming.

"Map - Reduce" "Hadoop"

