---
tags:
    - stat128
---

Functions

- Predict when variables will be available in function bodies and global environment
- Determine whether a function is vectorized.

Announcements:

- Recall late policy on assignments.
- Minor change to question on HW template just now

123 GO - What's the last book you read?

Assignment feedback

- Many discovered the formula interface for `plot(y ~ x, data = d)`. 👍
    Logical operators: `& |`
- Only necessary to convert date column one time at the beginning of your analysis.
    General principle: the less code, the better
- Summaries are fine, but don't print long vectors.
- Names should be semantically meaningful.
    Picking good names is really hard!
    descriptive: `roseville = air[..., ]`
    not descriptive: `data, mydata`
- Color should be semantically meaningful. We'll talk about color in graphs next week.
- `x == "TRUE"` What does this do? How about this? `x == TRUE`

```{r}
x = 1:10
smallx = x < 5
smallx
```

You should never need to do this for a logical vector.
```{r}
smallx == TRUE
```

What does this do?
Checks if smallx has the literal value "TRUE" by __coercing__ `smallx` to an object of class "character".
```{r}
smallx == "TRUE"
```

```{r}
smallx2 = as.character(smallx)  # explicit coercion
smallx2 == "TRUE"
```


```{r}
letters == "a"
```

## Function Scope

Define a function for 

$$
    f(x, a, y) = a * x + y
$$

```{r}
axpy = function(x, a = 1, y = 0)  # Arguments, including defaults
{   # <- Start of the body
    a * x + y           # Last element of body is the return value
}
```

123GO:

```{r}
axpy(x = 10, a = 5, y = 2)
```

What are the classes of the arguments that `axpy` accepts? Yes, numeric will work.
What are the names of the arguments that `axpy` accepts? `x, a, y`.

#### Argument Defaults

```{r}
axpy(x = 10, y = 2)
```

`a` defaulted to 1.

```{r}
axpy = function(x, a = 1, y = 0)
{
    print(a)
    a * x + y
}

axpy(10, y = 2)

# This will for a in global environment, and fail to find it.
#print(a) # 123GO - will this work or not? It's outside of my function
```


```{r}
head(trees)
```

`Girth` and `Height` make sense as variable names in the context of the `trees` data frame.
```{r}
with(trees, Girth + Height)
```


### Globals

```{r}

# z is global
z = 100

axpy = function(x, a = 1, y = 0)
{
    # x, a, y, are LOCAL to this function
    a * x + y + z
}

axpy(10)
```

```{r}
# Here air is a global variable.
plot_n_recent = function(d, n = 50)
{
    d2 = air[]    
}
```

Q: How can using global variables in our functions hurt us?

Easy to lose track of what's going on!

```{r}
#' Return the data point in `x` closest to the mean of `x`.
#'
closest_to_mean = function(x)
{   # x was precip, and I tried to change all appearance of precip to x
    diff_from_mean = abs(x - mean(x))
    closest_index = which.min(diff_from_mean)
    precip[closest_index]            # <--- FORGOT TO CHANGE TO X!!!!!! }
}
``` 

Using global variables in this way is one of the most common sources of errors.

How to avoid?

1. Specify and use default arguments, such as `n = 50`
2. Use only global functions in your function body. For example, `abs`
    Don't use data / actual values. Pass data in as arguments.
    
```{r}
axpy = function(x, a = 1, y = 0)
{
    a * x + y
}

axpy()
```


### Vectorized functions

Group activity: vectorized polynomial.

$$
f(x) = 2 * x^2 + 5*x + 100
$$
Vectorization means the function works elementwise.
It makes your code clean and fast.

```{r}
f = function(x) 2 * x^2 + 5*x + 100

#f(1)

#f(1:10)

x = 1:10
ifelse(x < 5, 100, -pi)
```





