

- Compare and contrast special values `Inf`, `NaN`, `NA`, `NULL`
- Predict propagation of `NA` through vectorized operations
- Test for presence of special values

Announcements:

- Visualization discussion assignment 
- Peer reviews for last homework
  - at least 1 positive comment, 1 idea for improvement
- Data science career panel coming up next Friday.

123 GO - What kind of data set sounds most interesting to you today?
  Air quality / fire, Politics, polls, police violence, finance, etc.

Note - Maybe finance and time series?

## `Inf`

`1/x`

```{r}
curve(1/x)
```

```{r}
1/0
```

> The basic rule should be that calls and relations with Infs really are statements with a proper mathematical limit.

123 GO
```{r}
Inf + 1
2 * Inf
-1 * Inf
mean(c(Inf, 1, 2, 3))
1/Inf
```

If a value is infinite, it can __propagate__ through further operations.

## `NaN`

```{r}
Inf / Inf
```

`NaN` stands for "Not a Number"

```{r}
0 / 0
```

```{r}
NaN + 1
2 * NaN
-1 * NaN
mean(c(NaN, 1, 2, 3))
1/NaN
```


## `NA`

```{r}
head(airquality)
```
`NA` means the value in a vector is missing.

```{r}
class(airquality$Ozone)
```

`NA` is R's "killer feature".
R handles `NA` in everything that it does, and this is very valuable for data analysis.

```{r}
NA + 1
2 * NA
-1 * NA
mean(c(NA, 1, 2, 3))
1/NA
```

`rm` in R means remove.

```{r}
mean(c(NA, 1, 2, 3), na.rm = TRUE)
```

```{r}
x = c(NA, 1, 2, 3)
x
```

Select values of `x` that are not NA
```{r}
is.na(x)
x[!is.na(x)]
mean(x[!is.na(x)])
```


## Imputing missing values

Imputing means to "fill in" missing values

```{r}
x[is.na(x)] = 0
x
```

An easy way to impute is to put the mean value in.

```{r}
a = airquality$Ozone
a[is.na(a)] = mean(a, na.rm = TRUE)
a
```


## `NULL`

What kind of object should `p` be?

```{r}
p = plot(1:10)
```

We need an object for a placeholder when we're not sure what the object should be.

```{r}
p
```

Why is `NULL` useful?

- Initiate variables / fill values
- Test for it, and handle it.

```{r}
is.null(p)
```



## Testing for special values

```{r}
x[1] = NA
x
```

What's wrong with this?

```{r}
x == 1

x == NA
```

```{r}
x[1] = NULL
x
```



## Other special values

```{r}
x = numeric()
x
length(x)
```

```{r}
s = ""
s

s2 = c("", "abc", "hello")
s2
```


## Summary


Sometimes arithmetic doesn't work out.
If this happens, then we might get one of the numeric special values.

- `Inf` means infinity
- `NaN` means "Not a Number"

There are two other more general special values.

- `NA` represents missing values in a vector, they usually show up when you load your data.
- `NULL` is a placeholder for ambiguous objects
