123 GO - what's your dream vehicle?

- Statistical models often try to estimate a response based on several predictors

Last homework:

- Going to change rubric
- General goal in statistics is to uncover general patterns and variance, and that's not really possible with just one observation.

## Statistical models

Often: want to estimate f, given observations of X and y

y ≅ f(X)

names for X - Predictors, variables, input

names for y - response, output, 

~~endogeneous exogeneous~~


## Interaction plots

In statistics, an __interaction effect__ is when two or more variables have a non additive effect in the value of some response.

One way to understand a model is to simulate it, so you can code in the exact assumptions of the model, and see what the results look like.
These have the advantage of being self contained too.

```{r}
#alcohol = 0:4
alcohol = c(3, 1, 0, 4, 2)
distractor = factor(c("None", "Hamburger", "Cell Phone"))
d = data.frame(alcohol = rep(alcohol, each = length(distractor)))
d$distractor = rep(distractor, times = length(alcohol))

d$prob_accident = 0
d$prob_accident = 0.02 * d$alcohol
cp = d$distractor == "Cell Phone"
d$prob_accident[cp] = d$prob_accident[cp] + 0.03

hb = d$distractor == "Hamburger"
d$prob_accident[hb] = d$prob_accident[hb] + 0.01

# Interaction effect between alcohol and cell phones
h34 = with(d, 3 <= alcohol & distractor == "Cell Phone")
d[h34, "prob_accident"] = d[h34, "prob_accident"] + 0.05

# Noise
set.seed(3810)
d$prob_accident = d$prob_accident + rnorm(nrow(d), sd = 0.008)

with(d,
interaction.plot(alcohol, distractor, prob_accident)
)
```

```{r}
library(ggplot2)
g = ggplot(data = d, aes(x = alcohol, y = prob_accident)) +
  geom_line(aes(color = distractor))
print(g)
```

Inputs: day of the week assignment is due, day of the week start assignment
Output: Grade

Inputs: alcohol, cigarette
Output: Life expectancy

Categorical ordered: fresh, soph, junior, senior
unordered: Age ranges: [18 - 27], 

Inputs: time spent outside, smoking, exposure to xrays, 
Output: Risk of cancer

Input: vacation time, type of job
Output: Job satisfaction

unordered categorical: gender
ordered categorical: intensity of exercise
output: weight loss

ordered: age ranges (0-9], (10-19], etc.
unordered: ethnicity
output: probability getting sick

## Office Hours

path separator `\` on Windows.
```{r}
```

s = "Hello
world
there"

```{r}
p = "E:/STAT128/128HW/COVID-19_Case_Surveillance_Public_Use_Data.csv"
```

For file paths on Windows you can use this slash: `/`

Find rows that actually contain `NA`

```{r}
cna = covid[!complete.cases(covid), ]

d = cna$death_yn
d == "Missing"
d
```

```{r}
table(covid$age_group, useNA = "always")
```

Apply the function table to each of the columns.

```{r}
lapply(covid[, c("age_group", "sex")], table, useNA = "always")
```

```{r}
t0 = table(covid$age_group)

# Because I don't know actual population:
total_pop = rep(0.1, times = 10)
total_pop[1] = 0.15
total_pop[9] = 0.05
total_pop = total_pop * 300e6

proportion_infected = t0 / total_pop

infected_per_100K = proportion_infected * 100e3
```

```{r}
count_miss = function(col_data)
{
  #keep = is.na(col_data) | col_data == "Unknown" | col_data == "Missing"
  count = 0
  for(i in seq_along(col_data))
  {
    if(is.na(col_data[i]))                           
      count = count + 1
    else {
      if(col_data[i] == "Unknown")
      count = count + 1
      if(col_data[i] == "Missing")
        count = count + 1
    }
  }
  count   
}
```


```{r}
system.time(count_miss(covid$age_group))
```


```{r}
count_miss(covid$age_group)
```

```{r}
system.time(table(covid$age_group, useNA = "always"))
```

Vectorized:
```{r}
count_miss_vec = function(col_data)
{
  keep = is.na(col_data) | col_data == "Unknown" | col_data == "Missing"
  sum(keep)
}
```


```{r}
system.time(count_miss_vec(covid$age_group))
```











