
rpart- recursive partitioning and regression trees

- Describe the rpart regression model at a high level
- Interpret a printed `rpart` object

123 GO – What are you looking forward to this weekend?

References:

"vignette" is a short narrative describing an R package

- [rpart vignette](https://cran.r-project.org/web/packages/rpart/vignettes/longintro.pdf)
- [regression trees](https://uc-r.github.io/regression_trees) from UCR


## objective function

Quick review:
The linear model says `y = ax + b + noise`.

It minimizes:
$$
\sum_{i = 1}^n (y_i - (ax_i + b))^2
$$

$$
\sum_{i = 1}^n (y_i - f(x_i))^2
$$
$$
\sum_{i = 1}^n (y_i - c)^2
$$

The `c = mean(y)`, sample mean minimizes this.
```{r}
library(rpart)
n = 100
x = sort(runif(n))
y = rnorm(n)
y[1:50] = y[1:50] + 20
plot(x, y)

fit = rpart(y ~ x)
lines(x, predict(fit))
```




## piecewise constant

```{r}
library(rpart)

n = 100

# 123 GO: Does it really matter? No, not really.
#x = sort(3 * runif(n))
x = seq(from = 0, to = 3, length.out = n)
y = rep(2, n)
y[x < 1] = 1
y[2 < x] = 3
noise = rnorm(n, sd = 0.2)
y = y + noise
d = data.frame(x, y)

plot(x, y)
fit = rpart(y~x)
lines(x, predict(fit))

```
```{r}
fit
```

This model is piecewise constant.

## rpart

## greedy algorithms

To see the parameters for `rpart`:

```{r}
?rpart.control
```

## group activity

```{r}
#' Calculate the mean squared error of the model on testdata
#'
#' @param model fitted model object
#' @param testdata data frame containing a column y representing the true values to compare the fitted values against
mse = function(model, testdata)
{
  yhat = predict(model, testdata) # Rely on predict to do everything specific to the model.
  y = testdata[, "y"]
  d2 = (yhat - y)^2
  mean(d2)
}
```


```{r}
# Simulate data
n = 200
d = data.frame(x = runif(n))
d$y = 10 * d$x + 2 + rnorm(n) # True model is a linear model

# Test / train split
test_index = sample(n, size = round(n/2))
dtrain = d[test_index, ]
dtest = d[-test_index, ]

fit = lm(y ~ x, data = dtrain)
fit2 = rpart(y ~ x, data = dtrain)
mse(fit, dtest)
mse(fit2, dtest)
```

