Goals:

- Utility scripting
- Read, write, rename, and move files in directories

123 GO – One good memory from this Thanksgiving

Announcements:

- Ask me in Discord what you want to cover more for these last two weeks
- Course evaluations (will talk about after class today)

## File and directory structure

Directories (folders) contain files.
123 GO: Give an example of a kind of file.

We're often interested in a handful of files:

- `.Rmd` files, R markdown documents
- `.html` files, reports generated from R markdown
- `.R` files, R scripts
- `.csv` files, containing tabular data

Much of data analysis is simply reading, writing, and organizing files.
Some people call this __utility scripting__.
It saves you the trouble of doing the same thing over and over again.
For example, resize every image in a directory.

Computers do repetition better than people do.
Anyone got any fun examples of doing the same thing over and over manually?

- manually checking prices for baseball cards, camelcamelcamel

Do the machine readable format first, then make the human version.


## Working directory

```{r}
getwd()
```

look at the files in this directory:
```{r}
list.files()
```
Change the directory:

```{r}
setwd("~")
```


## Making new directories.

```{r}
d = read.csv(text = "language,word
en,hello
en,goodbye
sp,hola
sp,adios
kor,안녕")
```

I want a directory called "languages".
Want to split this data, and save each language in its own file.

```{r}
dir.create("languages")
```


## Relative directories

Relative:

```{r}
list.files("languages")
```

Absolute:

```{r}
list.files("/Users/fitzgerald/projects/stat128/languages")
```

If you have a script to run possibly on different computers, should you use absolute or relative paths?
123 GO: Relative paths


## Writing the files

```{r}
en = d[d$language == "en", ]

en_file = paste0(en[1, "language"], ".csv")

en_file2 = file.path("languages", en_file)

write.csv(en, en_file2, row.names = FALSE)
```


Let's do all of them:

```{r}
en = d[d$language == "en", ]
sp = d[d$language == "sp", ]
kor = d[d$language == "kor", ]
...
```

violates DRY- don't repeat yourself.

Write a function:

```{r}
en = d[d$language == "en", ]

save_one_language = function(lang)
{
  lang_file = paste0(lang[1, "language"], ".csv")
  lang_file2 = file.path("languages", lang_file)
  write.csv(lang, lang_file2, row.names = FALSE)
}

save_one_language(en)
```

Apply to all languages:

```{r}
list_of_all_languages = split(d, d$language)
lapply(list_of_all_languages, save_one_language)
```


## Reading the files in

Goal is to reconstruct `d`.

Not ideal, violates DRY:

```{r}
en2 = read.csv("languages/en.csv")
kor2 = read.csv("languages/kor.csv")
sp2 = read.csv("languages/sp.csv")
```

```{r}
lang_files = list.files("languages", full.names = TRUE)

d2 = lapply(lang_files, read.csv)

d2 = do.call(rbind, d2)
```

Now we have `d`, up to row permutations.



## Cleaning up

```{r}
unlink("languages", recursive = TRUE)
```

BOOM.

```{r}

read.csv("vvre0519sft01.csv", skip = 10)

```
