Outcomes:

- Follow standard workflow for prediction problems
- Develop functions in the course of analysis
- Plot as you go to check assumptions

Plan: Demonstrate outline of solution to interest prediction assignment and address common mistakes.

123 GO: One thing you are thankful for.

Announcements:

- Course grades look high because you currently see full points for final project.
- Final project announcement to come soon.
- Discord - ask questions any time.


```{r}
library(rpart)
```



### Test train split

Idea: random partition


```{r}
set.seed(280)

debt = read.csv("~/data/ca_city_debt.csv")
test_frac = 0.2

n = nrow(debt)
test_index = sample(n, size = n * test_frac)

test = debt[test_index, ]
train = debt[-test_index, ]
```



### Preprocessing

Getting the data clean enough for us to use.

```{r}

refine = function(d)
{
# The column "interest" is my y to predict.
# Implicit in this coding is that TIC or NIC is **more** important.
# 123 GO: which one: TIC
d$interest = d[, "TIC.Interest.Rate"]
missing = is.na(d$interest)
d$interest[missing] = d[missing, "NIC.Interest.Rate"]

# New columns:
d$year = as.integer(substring(d$CDIAC.Number, 1, 4))
d$bond = grepl("bond", d$Debt.Type, ignore.case = TRUE)
mi = is.na(d$interest) 
message("Proportion of interest missing: ", mean(mi))
d[!mi, c("interest", "year", "bond")]
}

train2 = refine(train)
```

Quick visualization:

```{r}
boxplot(interest ~ bond, train2)

boxplot(interest ~ year, train2)
```


### Models

```{r}
fit_lm = lm(interest ~ ., train2)
```

```{r}
fit_lm
```

```{r}
fit_rpart = rpart(interest ~ ., train2)
```

```{r}
# library(plot.rpart)
plot(fit_rpart, xlim = c(0, 7), ylim = c(0.3, 1.5))
text(fit_rpart)
```


### Evaluating Performance

`fit_lm` has never seen the data in `test` or `test2`.

```{r}
test2 = refine(test)

test2$lm_pred = predict(fit_lm, test2)

# Error - how far away prediction was from the actual value
test2$difference = with(test2, interest - lm_pred)

hist(test2$difference)
```






