## Outcomes

- create logical vectors using comparison operators
- select rows by condition

Before we can do this, we need to know about dates:

- date class
- format strings
- conversion to dates
- compare dates
- update columns in place

### Announcements

- Today's online notes are more thorough, please use them in your homework.
- HW graded

123GO - what's your word of the day?

### References

- [Book of R, Section 5.2.3](https://learning.oreilly.com/library/view/the-book-of/9781492017486/xhtml/ch05.xhtml#ch05lev2sec54)
- [Art of R Programming, Chapter 5](https://learning.oreilly.com/library/view/the-art-of/9781593273842/ch05s02.html#extracting_subdata_frames)


## Dates

```{r}
air = read.csv("http://webpages.csus.edu/fitzgerald/files/stat128/fall20/ad_viz_plotval_data.csv")
```

In general, to select a subset of a data frame `data`
```
data[rows, columns]
```

Let's select columns

```{r}
d = air[, "Date"]
pm2 = air[, "Daily.Mean.PM2.5.Concentration"]
```

Vector or data.frame?

```{r}
class(d)
```

plot pm2 as a function of date

```{r}
plot(d, pm2)
```

What happened here?
We tried to plot strings and that doesn't make any sense.
strings AKA character vectors

```{r}
head(d)
```

Let's convert to dates.
```{r}
d2 = as.Date(d, format = "%m/%d/%Y")

class(d2)
```

123go - is it reasonable to plot dates on the x axis?
Yes, this is a time series.

```{r}
plot(d2, pm2)
```

side note:
function: input ----> output

Did `d` change?
No, it's still a character vector.
`air` didn't change either.

```{r}
class(air[, "Date"])
```

To update the Date column in `air`:

```{r}
air[, "Date"] = as.Date(d, format = "%m/%d/%Y")
```


## logical comparisons

Logical vectors are vectors where the value is either TRUE or FALSE (or NA, but that's for later).

```{r}
l1 = c(TRUE, TRUE, FALSE)
```

```{r}
x = c(1, 2, 2, 3, 18)
```

I want to check which elements of x are equal to 2.

```{r}
x == 2
```

To save this logical vector into a variable called `isx2`:
```{r}
isx2 = x == 2

isx2 <- x == 2
```

This means `=` (assignment) is TOTALLY different from `==` (equality comparison).

Your turn:
```{r}
x
x <= 2  # type in the chat, 123 go
```

```{r}
d3 = head(d2)
d3
```

To find the dates beyond January 12:
```{r}
as.Date("2020-01-12") < d3
```

To find the dates before January 12:

```{r}
as.Date("2020-01-12") > d3
```

## row selection

We can select subsets of the data that we're interested in.

```{r}
x = c(TRUE, TRUE, FALSE, TRUE)
x2 = c(TRUE, TRUE, TRUE, TRUE)

any(x)
any(x2)

all(x)
all(x2)
```


```
data[rows, columns]
```
We can take a logical subset of rows- sometimes called filtering / query

For example, to get the subset of rows that corresponds to "El Dorado" county.

Technique:
Build your query piece by pieces, such that every piece works and does what you expect.
Test each piece.

DO NOT: ~~just write something complicated and think it will work.~~

step 1: Select the COUNTY vector from air

```{r}
air[, "COUNTY"]

```
step 2: Create a logical vector identifying "El Dorado"

```{r}
air[, "COUNTY"] == "El Dorado"
```
step 3: Select the rows from `air` that correspond.

```{r}
edc = air[, "COUNTY"] == "El Dorado"
air2 = air[edc, ]
```

To test, I can look at all values of "COUNTY" in my `air2`.
```{r}
table(air2[, "COUNTY"])
unique(air2[, "COUNTY"])
```

If I don't want to use intermediate variable:

```{r}
rows = air[, "COUNTY"] == "El Dorado"
cols = c("COUNTY", "Date", "Daily.Mean.PM2.5.Concentration")
air4 = air[rows, cols]
air4
```

How many columns should `air4` have?




