
- Implement a function given a description of what it should do
- Describe R's formula syntax, `y ~ x`
- Select rows of a data frame by index

123GO - what's your dream job?

Announcements:

- For the homework, piece together parts from the lecture notes this week.
- The standards for homework grading are high, to help you grow.
- I'll provide an example of a submission that would earn a full score.

Warm call start:

```{r}
students = readLines("~/Desktop/roster_stat128.txt")
sample(students, 1)
```

## Defining Functions

Q: What is a (mathematical) function?
A: map from codomain ---> domain
A:          input    ---> exactly one output

Functions are the heart of R.

Find the data point closest to the mean.

```{r}
diff_from_mean = abs(precip - mean(precip))
closest_index = which.min(diff_from_mean)
precip[closest_index]
```

Decide exactly how it should work.

123GO come up with a simple example of inputs and outputs
input: c()
output:

```{r}
#' Find the data point closest to the mean.   # <-- Comment (optional)
#
closest_to_mean = function(x)                 # <-- name, arguments
{                                           # <--- start of the body
  print(class(x))
  diff_from_mean = abs(x - mean(x))
  closest_index = which.min(diff_from_mean)
  x[closest_index]                # <--- last line is the return value
}

farthest_to_mean = function(x)                 # <-- name, arguments
{                                           # <--- start of the body
  diff_from_mean = abs(x - mean(x))
  farthest_index = which.max(diff_from_mean)
  x[farthest_index]                # <--- last line is the return value
}
```


```{r}
closest_to_mean(precip)
closest_to_mean(c(0, 2.1, 4))    # 2.1 should be the output
```

SUCCESS!

```{r}
# Doesn't make sense.
closest_to_mean(letters)
```


## Model Formulas

`y ~ x` means "model y as a function of x"
~ tilde "till - duh"

```{r}
?formula
```


```{r}
air = read.csv("http://webpages.csus.edu/fitzgerald/files/stat128/fall20/ad_viz_plotval_data.csv")
```

`boxplot`
```{r}
boxplot(air[, "DAILY_AQI_VALUE"])
```

Let's look at groups by county.

```{r}
boxplot(DAILY_AQI_VALUE ~ COUNTY, data = air)
```

```{r}
dav = air[, "DAILY_AQI_VALUE"]

boxplot(dav)
```


```{r}
boxplot(DAILY_AQI_VALUE ~ COUNTY, data = air)
```

Suppose I want to have 2 boxplots. One for Sacramento county, and one for everything else. How do I do this?

```{r}
air[, "sac"] = air[, "COUNTY"] == "Sacramento"
# TRUE means the county is Sacramento

# Original: boxplot(DAILY_AQI_VALUE ~ COUNTY, data = air)
boxplot(DAILY_AQI_VALUE ~ sac, data = air, xlab = "Site is in Sacramento County")
```


## Selecting Rows by Index


`table` will count up all different values in a contingency table.
```{r}
table(air[, c("COUNTY", "sac")])
```


```{r}
r = sample(length(letters))
r
```


```{r}
letters[r]
```


We can reorder rows in our data set.

```{r}
plot(air[, "DAILY_AQI_VALUE"], type = "l")
```


```{r}
air2 = air[order(air[, "DAILY_AQI_VALUE"]), ]

plot(air2[, "DAILY_AQI_VALUE"], type = "l")
```


