---
title: Final
date: 16 December 2020
---

#### Instructions

- Answer the following questions, and show all your R code.
- Upload your submission to Canvas in nicely formatted HTML generated from Rstudio.
- Each question is worth 2 points, for 10 points total.


## 1

Define a function `plot_red` as a wrapper function to `plot` that changes the default color to red (`col = "red"`).

```{r}
#plot_red(1:10)
```


## 2

Model `y` below as a linear function of `x`.
Show that the model predicts 19.54463 when x is 20.

```{r}
x = 1:10
y = x + sin(x)
```


## 3

The following code creates a directory containing several CSV files.

```{r, echo=FALSE, results='hide',message=FALSE}
datadir = "letters_for_stat128_final"
dir.create(datadir)
lapply(sample(letters, 10), function(l){
           d = data.frame(letter = l, number = 1:3)
           fname = file.path(datadir, paste0(l, ".csv"))
           write.csv(d, fname, row.names = FALSE)
})
```

Read all of these files in and combine them into one single data frame, `d`.
If you do it correctly, `d` will have 30 rows.
You can do this idiomatically in three lines of code.

```{r, echo=FALSE, results='hide',message=FALSE}
# Cleaning up, leave this as the last line
unlink("letters_for_stat128_final", recursive = TRUE)
```


## 4

Extract the 10 digit phone numbers from the following character vector.

```{r}
s = c("My number is 593-461-0728"
      , "My address is 123 Fake Street, and my phone number is 578-163-4290, thanks."
      , "Either call 269-501-3748 or email foo@bar.com.")

goal = c("593-461-0728", "578-163-4290", "269-501-3748")
```


## 5

Define a new `print` method for objects of class `final` that accepts any arguments and prints out `We're done!`.
Hint: use `cat()`.

```{r}
s = "This is it?"
class(s) = c("final", class(s))

s
```

Here's the behavior you should see in the console after you define your method:

```
> s
We're done!
>
```
