---
title: Covid Cases
date: 11 October 2020
---

#### Outcomes

- Download and load public data sets
- Read and interpret metadata
- Evaluate general claims using data, in the presence of missing data
- Create and interpret interaction plots


#### Instructions

- Answer the following questions, and show all your R code.
- Upload your submission to Canvas in nicely formatted HTML generated from Rstudio.

Download the [COVID-19 Case Surveillance Public Use Data](https://data.cdc.gov/Case-Surveillance/COVID-19-Case-Surveillance-Public-Use-Data/vbim-akqf) from the US Center for Disease Control (CDC).
Click on the "Export" button in the top right corner next to download the actual data set.


# Questions

## 1 - Reading the Documentation

_a) Describe this data set at a high level.
    What does each row mean?_

TODO - your answer

_b) How many rows and columns do you expect to see based on the website?
Load the data into R.
Is this reflected in the data?_

```{r}
# TODO: load data and look at dimensions
```

TODO - your answer


## 2 - Missing Data
 
_There are few `NA` values in this data, but much of the data is missing, unknown, or blank.
Which columns have the highest proportion of missing data?
Why might these be missing?_

```{r}
# TODO
```

TODO - your answer


## 3 - Age Risk

_The [CDC states](https://www.cdc.gov/coronavirus/2019-ncov/need-extra-precautions/older-adults.html) "Among adults, the risk for severe illness from COVID-19 increases with age, with older adults at highest risk."
Does the data support this claim?_

```{r}
# TODO
```

TODO - your answer


## 4 - Medical Condition Risk

_The [CDC states](https://www.cdc.gov/coronavirus/2019-ncov/need-extra-precautions/people-with-medical-conditions.html) "People of any age with certain underlying medical conditions are at increased risk for severe illness from COVID-19"
Does the data support this claim?_

```{r}
# TODO
```

TODO - your answer


## 5 - Interaction Effects

_In statistics, an __interaction effect__ is when two or more variables have a non additive effect in the value of some response.
In the questions above, we explored the severity of COVID-19 (response) given variables age and medical condition.
Do age and medical condition have an interaction effect in the severity of COVID-19?_

```{r}
# TODO hint: Use interaction.plot
```

TODO - your answer


## 6 - Your Question

_Come up with your own question about COVID-19, and answer it using this data set.
You may want to explore other demographic columns: sex, and race / ethnicity.
You could also look at how the number of cases have changed over time using the report dates._

```{r}
# TODO
```

TODO - Question and answer