---
title: Roulette
date: 18 October 2020
---

#### Outcomes

- Define custom functions
- Split, apply, and combine data


#### Instructions

- Answer the following questions, and show all your R code.
- Upload your submission to Canvas in nicely formatted HTML generated from Rstudio.

You can use the functions we've developed in class for this 

```{r}
source("https://raw.githubusercontent.com/clarkfitzg/stat128/master/roulette.R")
```


# Questions

## 1 - Long Term Behavior

_Write a function to produce a detailed 

TODO - your answer

_b) How many rows and columns do you expect to see based on the website?
Load the data into R.
Is this reflected in the data?_

```{r}
# TODO: load data and look at dimensions
```

TODO - your answer


## 2 - Missing Data
 
_There are few `NA` values in this data, but much of the data is missing, unknown, or blank.
Which columns have the highest proportion of missing data?
Why might these be missing?_

```{r}
# TODO
```

TODO - your answer


## 3 - Age Risk

_The [CDC states](https://www.cdc.gov/coronavirus/2019-ncov/need-extra-precautions/older-adults.html) "Among adults, the risk for severe illness from COVID-19 increases with age, with older adults at highest risk."
Does the data support this claim?_

_Hint 1: It might be useful to think about these COVID illness rates by age after adjusting     for the relative sizes of each age group in the country.
You'll need to bring in another data source to do this._


```{r}
# TODO
```

TODO - your answer



## 4 - Medical Condition Risk

_The [CDC states](https://www.cdc.gov/coronavirus/2019-ncov/need-extra-precautions/people-with-medical-conditions.html) "People of any age with certain underlying medical conditions are at increased risk for severe illness from COVID-19"
Does the data support this claim?_

```{r}
# TODO
```

TODO - your answer


## 5 - Interaction Effects

_In statistics, an __interaction effect__ is when two or more variables have a non additive effect in the value of some response.
In the questions above, we explored the severity of COVID-19 (response) given variables age and medical condition.
Do age and medical condition have an interaction effect in the severity of COVID-19?_

```{r}
# TODO hint: Use interaction.plot
```

TODO - your answer


## 6 - Your Question

_Come up with your own question about COVID-19, and answer it using this data set.
You may want to explore other demographic columns: sex, and race / ethnicity.
You could also look at how the number of cases have changed over time using the report dates._

```{r}
# TODO
```

TODO - Question and answer
